---
title: "Resample in R"
author: "Phil Yue Wu"
date: "12/22/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

myPackages <- c("tidyverse", 
                "lubridate", 
                "zoo", 
                "xts", 
                "chron", 
                "data.table", 
                "scales",
                "DT",
                "fTrading", 
                "data.table", 
                "knitr", 
                "gridExtra", 
                "plotly", 
                "stringr", 
                "kableExtra", 
                "maptools" ,
                "ggmap",
                "ggplot2",
                "ggsn",
                "sf",
                "leaflet", 
                "fmsb", 
                "RColorBrewer",
                #"plyr", # this library clash with dplyr all the time. 
                "png")
#"sp" ,
#"rgdal",
#"raster", 
#"rasterVis" ,
#"rgeos")

#devtools::install_github("tidyverse/ggplot2")
library(dplyr)
required_packages <- function(x,y){
  for( i in x ){
    if( ! require( i , character.only = TRUE ) ){
      if( Sys.info()["sysname"] == "Linux" ){
        install.packages( i , repos=y , type="source" , quiet=TRUE , dependencies = TRUE , verbose = FALSE )
      } else {
        install.packages( i , repos=y , quiet=TRUE , dependencies = TRUE , verbose = FALSE )
      }
      require( i , character.only = TRUE, quietly = TRUE )
    }
  }
}

required_packages((myPackages),"http://cran.rstudio.com/")
```

# Note for future improvements: 

The "aggregate" function is found to be the reason why the newly created grid misaligns with the original grid. In addition, the "fact" parameter in the function does not produce a resolution that I wanted. The problem is illustrated as follows. 

```{r}
leaflet(width="300px", height="300px") %>% 
  addTiles(options = providerTileOptions(maxZoom = 20,minZoom= 20)) %>% 
  addPolygons(data=soton_grid_resize, color="red", weight = 1 ) %>% 
  addPolygons(data=soton_grid, fillOpacity = 0) 
```

The map above shows that the two layers of grids are not aligned. Disaggregate does not split existing rasters. The plan is to create a raster using desireable resolution and extent. 

```{r}
## obtain extent of soton_grid, which I want the new grid to align with. 
## Because the soton_grid has been converted to wgs 84, degree, the resolution cannot be defined by meters
## So the first is to convert soton_grid back to crs 27700
soton_grid %>% 
  st_transform(27700) %>% 
  extent() -> soton_ext

grid_new <- raster::raster(soton_ext, res=100, crs="+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.9996012717 +x_0=400000 +y_0=-100000 +ellps=airy +datum=OSGB36 +units=m +no_defs")

## Check how much cells on each direcion, horizontal, ncol, and vertical, nrow
#grid_new

## Assign a number to each cell, so the raster is not NULL valued
values(grid_new) <- 1:ncell(grid_new)

## Convert the raster to polygons and change to a CRS that leaflet recongnises (WSG 84)
grid_new <- grid_new %>% 
  rasterToPolygons() %>%  # result of this function is a st layer
  st_as_sf() %>%  # change to sf object
  st_transform(4326)


leaflet() %>% 
  addPolygons(data = grid_new) %>% 
  addPolygons(data = soton_grid, color="red", fillOpacity = 0)
```





